# pre-class


Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
glimpse(gapminder)
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
gapminder %>%
  distinct(country, continent)  %>%
      count(continent)

#I call the dataframe, get a list of distinct (i.e. unique) countries and their corresponding continents, and then count up how many fall into each continent
```
2. Which European nation had the lowest GDP per capita in 1997? 
```{r}
gapminder %>% 
  filter(continent == "Europe", year == 1997) %>%
         filter(min_rank(gdpPercap) <=5) %>%
          select(country)
#First I limit us to looking at European nations in 1997 with filter.  Next I filter for the 5 lowest ranked (i.e. lowest GDP per capita) countries. Finally I select to only get countries in the end.
      
```
3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  filter(year < 1990 & year >= 1980) %>%
    group_by(continent) %>%
      summarize(mean(lifeExp))
#I filter so we are looking in the 80s, group by continent and find the average for each continent
```
4. What 5 countries have the highest total GDP over all years combined?
```{r}
GDP <- gapminder %>% mutate(GDP = gdpPercap*pop) #calculate gdp from percap and pop
GDP.total <- GDP %>% group_by(country) %>% summarize(GDPtotal= sum(GDP)) #here I calculate the total GDP for each country over all of the years

GDP.total%>%
  filter(min_rank(GDPtotal) >=138) %>%
   select(country)  #here I  use the same machinery as in number 2, order the dataset, filter for the highest values in the variable of interest (here GDP.total) and finally, limit the output to countries
```
5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
gapminder %>%
  filter(lifeExp >=80) %>%
    select(country, lifeExp, year)

#straightforwardly, I filter for the condition and then select the desired rows to print in the appropriate order
```
6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}


#SAY WHAT?
```
7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
```{r}
#SAY WHAT?
```
8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
