# pre-class


Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
glimpse(gapminder)
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
gapminder %>%
  distinct(country, continent)  %>%
      count(continent)

#I call the dataframe, get a list of distinct (i.e. unique) countries and their corresponding continents, and then count up how many fall into each continent
```
2. Which European nation had the lowest GDP per capita in 1997? 
```{r}
gapminder %>% 
  filter(continent == "Europe", year == 1997) %>%
         filter(min_rank(gdpPercap) <=5) %>%
          select(country)
#First I limit us to looking at European nations in 1997 with filter.  Next I filter for the 5 lowest ranked (i.e. lowest GDP per capita) countries. Finally I select to only get countries in the end.
      
```
3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  filter(year < 1990 & year >= 1980) %>%
    group_by(continent) %>%
      summarize(mean(lifeExp))
#I filter so we are looking in the 80s, group by continent and find the average for each continent
```
4. What 5 countries have the highest total GDP over all years combined?
```{r}
GDP <- gapminder %>% mutate(GDP = gdpPercap*pop) #calculate gdp from percap and pop
GDP.total <- GDP %>% group_by(country) %>% summarize(GDPtotal= sum(GDP)) #here I calculate the total GDP for each country over all of the years

GDP.total%>%
  filter(min_rank(GDPtotal) >=138) %>%
   select(country)  #here I  use the same machinery as in number 2, order the dataset, filter for the highest values in the variable of interest (here GDP.total) and finally, limit the output to countries
```
5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
gapminder %>%
  filter(lifeExp >=80) %>%
    select(country, lifeExp, year)

#straightforwardly, I filter for the condition and then select the desired rows to print in the appropriate order
```
6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
gapminder %>%
  group_by(country) %>%
  summarize(abscorrelation = abs(cor(lifeExp, gdpPercap))) %>%
    filter(min_rank(abscorrelation) >=133)
#first I group the countries and then find the absolute value of the correlation between life expectency and GDP per capita (for each country). Finally we filter so that we only see the 10 countries with the highest value for abscorrelation
```
7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
```{r}
poptable <- gapminder %>%
  group_by(continent, year) %>%
    filter(continent != "Asia") %>%
      summarize(mean(pop))  #creating poptable which has the average population for each continent-year excluding Asia
names(poptable) <- c("continent", "year", "meanpop") #named the column with the average population "meanpop"
poptable %>%
  colnames() #checked this worked
sorted <- poptable[order(desc(poptable$meanpop)),] #sorted the table descending by average population
sorted #displaying result
```
8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
```{R}
standardframe <- gapminder %>%
  group_by(country) %>%
    summarize(deviation = sd(pop, na.rm =T)) #find the sd of each country
standardframe

standardframe %>%
  filter(min_rank(deviation) <= 3) #take the three smallest standard deviations and display the country and their sd
  
```  
9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
```{r}
gm1992 <- gapminder %>% filter(year == 1992)
gm1992
typeof(gm1992) #shows that it is a list
```
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
```{r}
lastanswer <- gapminder %>%
  sort(year) %>%
    group_by(country)
      filter(pop<lag(pop), lifeExp>lag(lifeExp)) #I arrange the dataset to be by increasing year and then within each country filter for years in which the population decreased and the life expectancy increased (by sorting by year we can use the lag fucntion to do this)
lastanswer
```
